---
title: "Hands-on_Ex06"
execute:  
  warning: false
date: "`r Sys.Date()`"
---

# **9. Global Measures of Spatial Autocorrelation**

## 9.0 Getting Started

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## **9.3.1 Import shapefile into r environment**

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **9.3.2 Import csv file into r environment**

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")

```

### **9.3.3 Performing relational join**

```{r}
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)

```

### **9.3.4 Visualising Regional Development Indicator**

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)


```

## **9.4 Global Measures of Spatial Autocorrelation**

### **9.4.1 Computing Contiguity Spatial Weights**

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

### **9.4.2 Row-standardised weights matrix**

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

## **9.5 Global Measures of Spatial Autocorrelation: Moran’s I**

### **9.5.1 Maron’s I test**

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

### **9.5.2 Computing Monte Carlo Moran’s I**

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

### **9.5.3 Visualising Monte Carlo Moran’s I**

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

## **9.6 Global Measures of Spatial Autocorrelation: Geary’s C**

### **9.6.1 Geary’s C test**

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

### **9.6.2 Computing Monte Carlo Geary’s C**

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

### **9.6.3 Visualising the Monte Carlo Geary’s C**

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## **9.7 Spatial Correlogram**

### **9.7.1 Compute Moran’s I correlogram**

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)

```

```{r}
print(MI_corr)
```

### **9.7.2 Compute Geary’s C correlogram and plot**

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

```{r}
print(GC_corr)
```

# **10  Local Measures of Spatial Autocorrelation**

### **10.2.3 Setting the Analytical Toolls**

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## **10.3 Getting the Data Into R Environment**

### **10.3.1 Import shapefile into r environment**

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **10.3.2 Import csv file into r environment**

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### **10.3.3 Performing relational join**

```{r}
hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

### **10.3.4 Visualising Regional Development Indicator**

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

## **10.4 Global Spatial Autocorrelation**

### **10.4.1 Computing Contiguity Spatial Weights**

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

### **10.4.2 Row-standardised weights matrix**

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

### **10.4.3 Global Spatial Autocorrelation: Moran’s I**

### **10.4.4 Maron’s I test**

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

#### 10.4.4.1 Computing Monte Carlo Moran’s I

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

#### 10.4.4.2 Visualising Monte Carlo Moran’s I

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

### **10.4.5 Global Spatial Autocorrelation: Geary’s**

#### 10.4.5.1 Geary’s C test

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

#### 10.4.5.2 Computing Monte Carlo Geary’s C

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

#### 10.4.5.3 Visualising the Monte Carlo Geary’s C

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## **10.5 Spatial Correlogram**

### **10.5.1 Compute Moran’s I correlogram**

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)

```

```{r}
print(MI_corr)
```

### **10.5.2 Compute Geary’s C correlogram and plot**

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

```{r}
print(GC_corr)
```

## **10.6 Cluster and Outlier Analysis**

### **10.6.1 Computing local Moran’s I**

```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

```{r}
printCoefmat(data.frame(
  localMI[fips,], 
  row.names=hunan$County[fips]),
  check.names=FALSE)
```

```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

#### 10.6.1.2 Mapping local Moran’s I values

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

```

#### 10.6.1.3 Mapping local Moran’s I p-values

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

#### 10.6.1.4 Mapping both local Moran’s I values and p-values

```{r}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

## **10.7 Creating a LISA Cluster Map**

### **10.7.1 Plotting Moran scatterplot**

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

### **10.7.2 Plotting Moran scatterplot with standardised variable**

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% 
  as.vector 
```

```{r}
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

### **10.7.3 Preparing LISA map classes**

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
```

```{r}
hunan$lag_GDPPC <- lag.listw(rswm_q, hunan$GDPPC)
DV <- hunan$lag_GDPPC - mean(hunan$lag_GDPPC)     
```

```{r}
LM_I <- localMI[,1] - mean(localMI[,1])    
```

```{r}
signif <- 0.05       
```

```{r}
quadrant[DV <0 & LM_I>0] <- 1
quadrant[DV >0 & LM_I<0] <- 2
quadrant[DV <0 & LM_I<0] <- 3  
quadrant[DV >0 & LM_I>0] <- 4      
```

```{r}
quadrant[localMI[,5]>signif] <- 0
```

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
hunan$lag_GDPPC <- lag.listw(rswm_q, hunan$GDPPC)
DV <- hunan$lag_GDPPC - mean(hunan$lag_GDPPC)     
LM_I <- localMI[,1]   
signif <- 0.05       
quadrant[DV <0 & LM_I>0] <- 1
quadrant[DV >0 & LM_I<0] <- 2
quadrant[DV <0 & LM_I<0] <- 3  
quadrant[DV >0 & LM_I>0] <- 4    
quadrant[localMI[,5]>signif] <- 0
```

### **10.7.4 Plotting LISA map**

```{r}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```

```{r}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, 
             asp=1, ncol=2)
```

## **10.8 Hot Spot and Cold Spot Area Analysis**

### **10.8.1 Getis and Ord’s G-Statistics**

The analysis consists of three steps:

-   Deriving spatial weight matrix

-   Computing Gi statistics

-   Mapping Gi statistics

### **10.8.2 Deriving distance-based weight matrix**

#### 10.8.2.1 Deriving the centroid

```{R}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

```{r}
coords <- cbind(longitude, latitude)
```

#### 10.8.2.2 Determine the cut-off distance

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

#### 10.8.2.3 Computing fixed distance weight matrix

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

```{r}
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

### **10.8.3 Computing adaptive distance weight matrix**

```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn
```

```{r}
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)

```

## **10.9 Computing Gi statistics**

### **10.9.1 Gi statistics using fixed distance**

```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
gi.fixed
```

```{r}
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```

### **10.9.2 Mapping Gi values with fixed distance weights**

```{r}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

### **10.9.3 Gi statistics using adaptive distance**

```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
```

### **10.9.4 Mapping Gi values with adaptive distance weights**

```{r}
gdppc<- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) + 
  tm_fill(col = "gstat_adaptive", 
          style = "pretty", 
          palette="-RdBu", 
          title = "local Gi") + 
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, 
             Gimap, 
             asp=1, 
             ncol=2)

```
