---
title: "In-Class Exercise 2"
format: 
  html: 
    code-fold: false 
    code-summary: "Show the code"
execute:
  eval: true  #display the code chunk 
  echo: true 
  warning: false
date: "`r Sys.Date()`"
---

# Getting Started

In this hands-on exercise, the following R packages will be used:

-   arrow

-   lubridate

-   tidyverse

-   sf

-   tmap

```{r}
pacman::p_load(arrow, lubridate, tidyverse, sf, tmap)
```

## Importing Grab-Posisi Dataset

```{r}
df <- read_parquet("data/GrabPosisi/part-00000-8bbff892-97d2-4011-9961-703e38972569.c000.snappy.parquet")
```

Convert a code chunk to convert the data type of pingtimestamp from character to date-time

```{r}
df$pingtimstamp <- as_datetime(df$pingtimestamp)

```

```{r}
##write_rds(df, "data/rds/part0.rds")
```

# Extracting trip starting location

Using the steps we learnt before

-   extract trip origin locations

-   derive 3 new columns (i.e. variables) for weekdays, starting hour and day of the month

-   name the output tibble data.frame origin_df

```{r}
origin_df <- df %>% 
  group_by(trj_id) %>%
  arrange(pingtimestamp) %>%
  filter(row_number()==1) %>%
  mutate(weekday = wday(pingtimstamp,
                        label = TRUE,
                        abbr = TRUE),
         start_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

# Extracting trip ending locations

```{r}
destination_df <- df %>%
  group_by(trj_id) %>%
  arrange(desc(pingtimestamp)) %>% 
  filter(row_number()==1) %>% 
  mutate(weekday = wday(pingtimestamp,
                        label = TRUE,
                        abbr = TRUE),
         end_hr = factor(hour(pingtimestamp)),
         day = factor(mday(pingtimestamp)))
```

#writing rds file

```{r}
write_rds(origin_df, "data/rds/origin_df.rds")
write_rds(destination_df, "data/rds/destination_df.rds")
```

#Importing file (the next time we can use this directly)

```{r}
origin_df <- read_rds("data/rds/origin_df.rds")
destination_df <- read_rds("data/rds/destination_df.rds")
```

```{r}
read_rds("data/rds/origin_df.rds")
```
