---
title: "Hands-on_Ex11: 16  Calibrating Spatial Interaction Models with R"
format: html
date-modified: "last-modified"
execute:  
  warning: false
  freeze: true
  message: false 
  eval: true 
  echo: true 
  fig-retine: 3
date: "`r Sys.Date()`"
---

## **16.3 Getting Started**

```{r}
pacman::p_load(tmap, sf, sp,
               performance, reshape2,
               ggpubr, tidyverse)
```

## **16.4 The Data**

-   *od_data.rds*, weekday morning peak passenger flows at planning subzone level.

-   *mpsz.rds*, URA Master Plan 2019 Planning Subzone boundary in simple feature tibble data frame format.

## **16.5 Computing Distance Matrix**

```{r}
mpsz <- read_rds("data/rds/mpsz.rds")
mpsz
```

### **16.5.1 Converting from sf data.table to SpatialPolygonsDataFrame**

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

### **16.5.2 Computing the distance matrix**

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
```

```{r}
head(dist, n=c(10, 10))
```

### **16.5.3 Labelling column and row heanders of a distance matrix**

```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

### **16.5.4 Pivoting distance value by SUBZONE_C**

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```

### **16.5.5 Updating intra-zonal distances**

```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()
```

```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

```{r}
distPair %>%
  summary()
```

```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```

```{r}
write_rds(distPair, "data/rds/distPair.rds") 
```

```{r}
distPair <- read_rds("data/rds/distPair.rds")
```

## **16.6 Preparing flow data**

```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```

```{r}
flow_data <- od_data_fii %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

```{r}
head(flow_data, 10)
```

### **16.6.1 Separating intra-flow from passenger volume df**

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```

### **16.6.2 Combining passenger volume data with distance value**

```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```

## **16.7 Preparing Origin and Destination Attributes**

### **16.7.1 Importing population data**

```{r}
pop <- read_csv("data/aspatial/pop.csv")
```

### **16.7.2 Geospatial data wrangling**

```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

### **16.7.3 Preparing origin attribute**

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

### **16.7.4 Preparing destination attribute**

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

```{r}
write_rds(flow_data1, "data/rds/SIM_data.rds")
```

## **16.8 Calibrating Spatial Interaction Models**

### **16.8.1 Importing the modelling data**

```{r}
SIM_data <- read_rds("data/rds/SIM_data.rds")
```

### **16.8.2 Visualising the dependent variable**

```{r}
ggplot(data = SIM_data,
       aes(x = TRIPS)) +
  geom_histogram()
```

```{r}
ggplot(data = SIM_data,
       aes(x = dist,
           y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ggplot(data = SIM_data,
       aes(x = log(dist),
           y = log(TRIPS))) +
  geom_point() +
  geom_smooth(method = lm)
```

### **16.8.3 Checking for variables with zero values**

```{r}
summary(SIM_data)
```

```{r}
SIM_data$DESTIN_AGE7_12 <- ifelse(
  SIM_data$DESTIN_AGE7_12 == 0,
  0.99, SIM_data$DESTIN_AGE7_12)
SIM_data$DESTIN_AGE13_24 <- ifelse(
  SIM_data$DESTIN_AGE13_24 == 0,
  0.99, SIM_data$DESTIN_AGE13_24)
SIM_data$DESTIN_AGE25_64 <- ifelse(
  SIM_data$DESTIN_AGE25_64 == 0,
  0.99, SIM_data$DESTIN_AGE25_64)
SIM_data$ORIGIN_AGE7_12 <- ifelse(
  SIM_data$ORIGIN_AGE7_12 == 0,
  0.99, SIM_data$ORIGIN_AGE7_12)
SIM_data$ORIGIN_AGE13_24 <- ifelse(
  SIM_data$ORIGIN_AGE13_24 == 0,
  0.99, SIM_data$ORIGIN_AGE13_24)
SIM_data$ORIGIN_AGE25_64 <- ifelse(
  SIM_data$ORIGIN_AGE25_64 == 0,
  0.99, SIM_data$ORIGIN_AGE25_64)
```

```{r}
summary(SIM_data)
```

### **16.8.4 Unconstrained Spatial Interaction Model**

```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(ORIGIN_AGE25_64) + 
                log(DESTIN_AGE25_64) +
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
uncSIM
```

### **16.8.5 R-squared function**

```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

### **16.8.6 Origin (Production) constrained SIM**

```{r}
orcSIM <- glm(formula = TRIPS ~ 
                 ORIGIN_SZ +
                 log(DESTIN_AGE25_64) +
                 log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(orcSIM)
```

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)
```

### **16.8.7 Destination constrained**

```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(ORIGIN_AGE25_64) + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(decSIM)
```

```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

**16.8.8 Doubly constrained**

```{r}
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_SZ + 
                DESTIN_SZ + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(dbcSIM)
```

```{r}
CalcRSquared(dbcSIM$data$TRIPS, dbcSIM$fitted.values)
```

### **16.8.9 Model comparison**

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```

### **16.8.10 Visualising fitted values**

```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM$fitted.values")
```

```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM$fitted.values")
```

```{r}
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(decTRIPS = "decSIM$fitted.values")
```

```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM$fitted.values")
```

```{r}
unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```

```{r}

```
